### 2021-04-18 old/old_isipta.tex
Initial Overleaf Import

\documentclass[pmlr,twocolumn,a4paper,10pt]{jmlr}
% \documentclass[pmlr,twocolumn,10pt]{jmlr} % useful when printing your own Letter copy; should not impact layout; do not use for submission
% This class automatically loads the following packages:
% amsmath, amssymb, natbib, graphicx, xcolor, url, algorithm2e, hyperref
  \citestyle{plain} % comment to get author-year; see the natbib docs for more options
\usepackage{isipta2021}
% This style file
% * sets the inputencoding to utf8 (which should be used),
% * loads mathtools (fixes for amsmath) and
% * the font packages textcomp, mathptmx, helvet, and courier (so no font packages should be loaded) and neither should font size be changed
%
%% Choose your variant of English; be consistent
\usepackage[american]{babel}
\usepackage{algorithm,algorithmic,bm}
% \usepackage[british]{babel}
%% Suggested packages, as needed:
% \usepackage{siunitx} % for proper typesetting of numbers and units
\usepackage{booktabs} % commands to create good-looking tables
% \usepackage{rotating} % for sideways figures and tables
% \usepackage{longtable} % for long tables
\usepackage{tikz} % nice language for creating drawings and diagrams
%% Self-defined macros
\newcommand{\swap}[3][-]{#3#1#2} % just an example
%% Frontmatter
% Authors: DO NOT EDIT
\jmlrworkshop{ISIPTA 2021}
\title[Short Title]{% [Short title] can be omitted if title is short enough
  Credal Adaptive Testing
  %Title in Title Case;\titlebreak~ % tilde needed to get space in short title
  %Use the Same Style for Your Section Titles
}

\author{% Authors are listed one per line and grouped per affiliation
  \Name{permutation of Claudio/Giorgia/Francesca/Alessandro}\\ % Accented characters require some attention
  \addr IDSIA
  %\AND
  %\Name{Andrés Cano}\Email{acu@decsai.ugr.es}\\
  %  \Name{Serafín Moral}\Email{smc@decsai.ugr.es}\\
  %\addr Computer Science and Artificial Intelligence, University of Granada, Spain
  %\AND
  %\Name{Jasper {De Bock}}\Email{Jasper.DeBock@UGent.be}\\
  %\addr Electronics and Information Systems, Ghent University, Belgium
}

\begin{document}
\maketitle

\begin{abstract}% No length restrictions, but keep things reasonable
An \emph{adaptive} test is a computer-based testing technique which adjusts the sequence of questions on the basis of the estimated ability level of the test taker. We suggest the use of \emph{credal networks}, a generalization of Bayesian networks based on sets of probability mass functions, to implement adaptive tests exploiting the knowledge of the test developer instead of training on databases of answers. Compared to Bayesian networks, these models might offer higher expressiveness and hence a more reliable modeling of the qualitative expert knowledge. The counterpart is a less straightforward identification of the information-theoretic measure controlling the question-selection and the test-stopping criteria. We elaborate on these issues and propose a sound and computationally feasible procedure. Validation against a Bayesian-network approach on a benchmark about German language proficiency assessments suggests that credal networks can be reliable in assessing the student level and  effective in reducing the number of questions required to do it.
\end{abstract}
\begin{keywords}% Keywords provide context for readers, so put in up to about half a dozen
  keyword, mot clé, Williams-coherence, trefwoord, imprecise Markov chain
\end{keywords}

\section{Single Skill Model}
Let us introduce the discussion by means of a simple example modelling the knowledge level of a student in a particular domain. In order to evaluate such level a number of questions can be asked to the student. We describe the knowledge level (or skill) of the student by means of a single variable $S$. We assume $S$ to be a discrete, ordinal, variable, whose $m$ states $\mathcal{S}:=(s^{(0)},\ldots,s^{(m-1)})$ can be intended as increasing knowledge levels in that domain. Regarding the questions, we initially describe them as Boolean variables, whose true state correspond to a correct answer to the question, while no distinctions between the possible wrong answers are considered.\footnote{Dedicated discussion about abstention.} Let $Q$ with $\mathcal{Q}:=(q,\overline{q})$ denote a generic question we ask to the student. A repository of $n$ questions $\bm{Q}$ is assumed to be available. A \emph{testing} algorithm is a procedure deciding which question and how many questions from $\bm{Q}$ should be asked to the student, on the basis of her previous answers. Algorithm X provides an abstract description of xxx.

\begin{algorithm}[htp!]
    \caption{Given a dataset of student profiles $\bm{S}$\label{alg:simple}}
    \begin{algorithmic}[1]
        \FOR {$s\in\bm{S}$}
        \WHILE{ $h > h^*$}
        \STATE $Q^* \gets \mathrm{FindBestQuestion}(\bm{Q}$)
        \STATE $q^* \gets \mathrm{SimulateAnswer}(\sigma)$
        \STATE $\bm{e} \gets \bm{e} \cup \{ Q^*=q^* \}$
        \STATE $h \gets H[P(S|\bm{e}]$
        \ENDWHILE
        \ENDFOR
        \STATE {\bf return} $\arg\max P(s|\bm{e})$
    \end{algorithmic}
\end{algorithm}

\newpage

Let $S$ denote an ordinal variable with $m$ states, denoting the knowledge level of a student with respect to a particular knowledge domain. Let $Q$ denote a question to be

\section{Brainstorming}

\subsection{Linear Programs for Credal Mode}

Given the Boolean question $Q$ and the Boolean skill $S$, we use the score:

\begin{equation}
\sigma_Q := \max_{P\in K} \left[ 
\max \{ P(S=0,Q=0) , P(S=1,Q=0) \}  + \max \{ P(S=0,Q=1),P(S=1,Q=1)\}
\right]
\end{equation}

\begin{theorem}
\begin{equation}
\max_{x}  \max \{ f(x) , g(x) \}  
=
 \max \{ \max_{x'} f(x') , \max_{x''} g(x'') \} 
\end{equation}
Let 
$$y^{*}:=\max_x \max\{f(x),g(x)\}$$
$$y^{**}:= \max \{ \max_{x'} f(x') , \max_{x''} g(x'') \}$$
Say for instance that $y^{**}=max_x f(x)$ and let $x^{**}:=\arg\max_x f(x)$. Thus:
$$f(x^{**})\geq \max_{x'} g(x) \geq g(x^{**})$$

Let $x^{*}:= \arg\max_x \max\{f(x),g(x)\}$. 
It cannot be $y^{*}:=g(x^{*})$ as this would imply $g(x^{*}) \geq f(x^{*})$

%Assume ad absurdum $y^{*}\neq y^{**}$.

If $y^{*}:=f(x^{*})$ we should have $y^{*}=y^{**}$.
Vice versa,

or
%$y^{*}:=g(x^{*})$. In the first case 
\end{theorem}

\subsection*{Pseudocode}
Let us first specify the inputs:
\begin{itemize}
\item Let $\bm{S}:=(S_1,\ldots,S_N)$ denote the set of skills in the model. We describe a \emph{student} as a joint observation $\bm{s}$ of the skills.
\item Let $M$ denote the network \emph{skeleton}, i.e., the model over the skills only
\item A question $Q$ is a binary variables with states right and wrong
\item Each question depends on a single skill only, let $\mathrm{Pa}(Q)$ denote this skill
\item Let $\tau(S)$ denote the number of \emph{templates} for the CPT of a question associated with $S$, $\{P(Q|S)\}_{t=1}^{\tau(S)}$ 
\item Let $n_q$ be the maximum number of questions we might ask for each skill
\end{itemize}
The algorithm works as follows.
\begin{enumerate}
\item For eachLet $\bm{S}:=(S_1,\ldots,S_N)$ denote the set of skills in the model
\item Let $M$ denote the network \emph{skeleton}, i.e., the model over the skills only
\item A question $Q$ is a binary variables with states right and wrong
\item Each question depends on a single skill only, let $\mathrm{Pa}(Q)$ denote this skill
\item Let $\tau(S)$ denote the number of \emph{templates} for the CPT of a question associated with $S$, $\{P(Q|S)\}_{t=1}^{\tau(S)}$ 
\item Let $n_q$ be the maximum number of questions we might ask for each skill
\end{enumerate}

\begin{algorithm}[htp!]
    \caption{Single skill, precise model\\
    a model (BN, skeleton) $M$ over the skills only
    \\a dataset of $n$ students $\{s^i\}_{i=1}^n$,\\a dataset of $m$ questions $\mathcal{Q}:=\{Q^j\}_{j=1}^m$,\\
    the CPTs of the questions in $\mathcal{Q}$, i.e., $\{P(Q^j|S)\}_{j=1}^m$\\
    the student answer models $\{\tilde{P}(Q^j|S)\}$\\
    an entropy threshold $h^*$,\\
    a score function for questions $\gamma$
    \label{alg:simple}}
    \begin{algorithmic}[1]
    \FOR {$i\gets\{1,\ldots,n\}$}
    \STATE $(M',\mathcal{Q}',\mathcal{Q}'',\bm{e}) \gets (M,\mathcal{Q},\emptyset,\emptyset)$
    \WHILE{ $h > h^* \mathrm{\bf and}\, \mathcal{Q}' \neq \mathcal{Q}$}
    \STATE $Q^* \gets \arg\max_{Q \in \mathcal{Q}'}\gamma(M',Q,\bm{e})$
    \STATE{$\mathcal{Q}'\gets \mathcal{Q}' \setminus \{ Q^*\}$}
    \STATE{$\mathcal{Q}''\gets \mathcal{Q}'' \cup \{ Q^*\}$}
    \STATE{$q^* \gets \mathrm{sampling} \sim \tilde{P}(Q^*|s^i)$}
    \STATE $M' \gets \mathrm{addChild}(M',Q,S,P(Q|S))$
    \STATE $\bm{e} \gets \bm{e} \cup \{ Q^*=q^* \}$
    \STATE $h \gets H[P_{M'}(S|\bm{e})]$
    \ENDWHILE
    \STATE $\sigma_i \gets \arg\max_{s\in\mathcal{S}} P_{M'}(s|\bm{e})$
    \ENDFOR
    \STATE {\bf return} $(\sigma_1,\ldots,\sigma_n)$
    \end{algorithmic}
\end{algorithm}
Three possible scores:
\begin{eqnarray}
    \gamma(Q,\bm{e}) := -H[S|Q,\bm{e}] = -\sum_q P(q) \cdot H[S|\bm{e},q]&&\\
    \gamma'(Q,\bm{e}) :=  -\sum_q P(q) \cdot \max_{s\in\mathcal{S}}  P[s|\bm{e},q]&&\\
    \gamma''(Q,\bm{e}) :=  |P(Q=1|\bm{e})-0.5|&&
\end{eqnarray}

\textcolor{red}{Maybe, instead of (3) I would use a known measure of qualitative variation (\url{https://en.wikipedia.org/wiki/Qualitative_variation}) which extend easily to non binary answers, e.g., the variation ratio ($1-\max_q(P(q))$) or the M1 Gibb's index also known as coefficient of unalikeability and somehow related to the variance (it is the variance in the binary case): $M1 = 1-\sum_q P(q)^2$, or even the entropy, which is not so difficult to compute for $P(Q)$. They are all equivalent in the binary case, as they are all monotonic function of $\max_q(P(q))$ only. }

\begin{algorithm}[htp!]
    \caption{Single skill, credal model\\a dataset of $n$ students $\{s^i\}_{i=1}^n$,\\a dataset of $m$ questions $\mathcal{Q}:=\{Q^j\}_{j=1}^m$,\\
    the credal CPTs of the questions in $\mathcal{Q}$, i.e., $\{\overline{P}(Q^j|S)\}_{j=1}^m$\\
    an upper entropy threshold $\overline{h}^*$,\\
    a score function for questions $\sigma$
    \label{alg:simple}}
    \begin{algorithmic}[1]
    \FOR {$i\gets\{1,\ldots,n\}$}
    \STATE $(\mathcal{Q}',\mathcal{Q}'',\bm{e}) \gets (\mathcal{Q},\emptyset,\emptyset)$
    \WHILE{ $h > h^* \mathrm{\bf and}\, \mathcal{Q}' \neq \mathcal{Q}$}
    \STATE $Q^* \gets \arg\max_{Q \in \mathcal{Q}'}\gamma(Q,\bm{e})$
    \STATE{$\mathcal{Q}'\gets \mathcal{Q}' \setminus \{ Q^*\}$}
    \STATE{$\mathcal{Q}''\gets \mathcal{Q}'' \cup \{ Q^*\}$}
    \STATE{$q^* \gets \mathrm{sampling} \sim P(Q^*|s^i)$}
    \STATE $\mathrm{addChild}(Q,S,P(Q|S))$
    \STATE $\bm{e} \gets \bm{e} \cup \{ Q^*=q^* \}$
    \STATE $h \gets H[P(S|\bm{e})]$
    \ENDWHILE
    \STATE $\sigma_i \gets \arg\max_{s\in\mathcal{S}} P(s|\bm{e})$
    \ENDFOR
    \STATE {\bf return} $(\sigma_1,\ldots,\sigma_n)$
    \end{algorithmic}
\end{algorithm}
Three possible scores:
\begin{eqnarray}
    \gamma(Q,\bm{e}) := -H[S|Q,\bm{e}] = -\sum_q P(q) \cdot H[S|\bm{e},q]&&\\
    \gamma'(Q,\bm{e}) :=  -\sum_q P(q) \cdot \max_{s\in\mathcal{S}}  P[s|\bm{e},q]&&\\
    \gamma''(Q,\bm{e}) :=  |P(Q=1|\bm{e})-0.5|&&
\end{eqnarray}

\begin{algorithm}[htp!]
    \caption{Given a dataset of student profiles $\bm{S}$\label{alg:simple}}
    \begin{algorithmic}[1]
        \FOR {$s\in\bm{S}$}
        \WHILE{ $h > h^*$}
        \STATE $Q^* \gets \mathrm{FindBestQuestion}(\bm{Q}$)
        \STATE $q^* \gets \mathrm{SimulateAnswer}(\sigma)$
        \STATE $\bm{e} \gets \bm{e} \cup \{ Q^*=q^* \}$
        \STATE $h \gets H[P(S|\bm{e}]$
        \ENDWHILE
        \ENDFOR
        \STATE {\bf return} $\arg\max P(s|\bm{e})$
    \end{algorithmic}
\end{algorithm}

\subsection*{Formulae}
Let us define the functions, all defined for $x\in [0,1]$, apart from $f$ defines for each $x\geq 0$:
\begin{eqnarray}
f(x)&:=& \frac{1}{1+x},\,,\\
h(x)&:=&-x \cdot \log_2 x - (1-x)  \cdot \log_2 (1-x)\,,\\
l(x)&:=&\frac{x}{1-x}\,,\\
m(x)&:=&2 \max\{x,1-x\}\,,\\
i_\epsilon(x)&:=&\epsilon\log(e^{\frac{2x}{\epsilon}}+e^{\frac{2(1-x)}{\epsilon}})\,,\\
\tilde{i}_\epsilon(x)&:=&\frac{i_\epsilon(x)-i_\epsilon(0)}{i_\epsilon(\frac{1}{2})-i_\epsilon(0)}\,.
\end{eqnarray}

Derivatives:
\begin{eqnarray}
f'(x)&:=& -\frac{1}{(1+x)^2},\,,\\
h'(x)&:=&\log_2 \frac{1-x}{x}\,\\
G(x) &:=& h \circ f(x)\,\\
G'(x) &:=& h[f(x)] f'(x)=-\frac{\log_2\frac{1-f(x)}{f(x)}}{(1+x)^2}\,\\
G'(x) &:=& -\frac{\log_2 x}{(1+x)^2}\,\\
l(x)&:=&\frac{x}{1-x}\,,\\
m(x)&:=&2 \max\{x,1-x\}\,,\\
i_\epsilon(x)&:=&\epsilon\log(e^{\frac{2x}{\epsilon}}+e^{\frac{2(1-x)}{\epsilon}})\,,\\
\tilde{i}_\epsilon(x)&:=&\frac{i_\epsilon(x)-i_\epsilon(0)}{i_\epsilon(\frac{1}{2})-i_\epsilon(0)}\,.
\end{eqnarray}

A few remarks. Function $f$ is a monotonically decreasing function with values in $[0,1]$. Function $h(x)$ corresponds to the entropy of a mass function over a Boolean variable with a probability equal to $x$. On the extreme points of its domain, we take the limit value, i.e., zero. The function is unimodal and achieve its maximum on  $x=\frac{1}{2}$. Function $l$ is the popular \emph{logit} function, it is monotonically increasing function of $x$ and $l(1-x)=l(x)^-1$. Function $m(x)$ is $x$ for $x\geq \frac{1}{2}$ and $1-x$ otherwise. 
Exactly as $h$, functions $m$ and $\tilde{i}_\epsilon$, for each $\epsilon>0$, have the same (zero) value for $x=0$ and $x=1$, being all unimodal with maximum in $x=\frac{1}{2}$.
Finally, the popular Gumbel's trick implies:
\begin{equation}
\lim_{\epsilon \to 0} i_\epsilon(x)= 
\lim_{\epsilon \to 0} \tilde{i}_\epsilon(x)
= m(x)\,.
\end{equation}
\subsection*{Another Formula}
Consider a PMF $P(X)$. Let $x^*:=\arg\max_{x\in\mathcal{X}} P(x)$. Let $\mathcal{P}$ denote the set of all PMFs with the same $p^*:=P(x^*)$. The most entropic PMF in $\mathcal{P}$ is the one with  $P(x)=\frac{1-p^*}{m-1}$ for each $x\neq x^*$ where $m:=|\mathcal{X}|$. While the least entropic is the one with a $x'\in\mathcal{X}\setminus\{x^*\}$ such that $P(x')=1-p^*$ and all the other masses equal to zero. These two bounds on the entropy are therefore:
\begin{eqnarray}
\overline{h}&:=& -p^* \log_m p^* - (1-p^*) \log_m \frac{1-p^*}{m-1}\,,\\
\overline{h}&:=& -p^* \log_m p^* - (1-p^*) \log_m (1-p^*)\,.
\end{eqnarray}
This:
\begin{equation}
\overline{h}-\underline{h}=(1-p^*) \log_m (m-1)\,,
\end{equation}
The maximum gap between the entropies of the PMFs in $\mathcal{P}$ decreases as soon as $p^*$ increases and as soon as the number of states of $X$ decreases, being zero for binary variables.

\subsection*{Toy Example}
Consider a Boolean skill $S$ with states $s_0$ and $s_1$ and a Boolean question/answer $Q$. Let us shape the model as a BN, $S\to Q$. The prior $P(S)$ is specified by a parameter $p:=P(s_1)$. The question CPT $P(Q|S)$ might be specified instead by two parameters $\pi_1:=P(q_1|s_1)$ and $\pi_0:=P(q_1|s_0)$. The informativeness score we assign two the question is
\begin{equation}
\sigma_Q = H[S|Q]:=H[S|q_0]P(q_0)+H[S|q_1]P(q_1)\,.
\end{equation}
We have
\begin{equation}
P(q_1)=\pi_1 p + \pi_0 (1-p)
\end{equation}
while $P(q_0)=1-P(q_1)$ and also
\begin{equation}
P(s_1|q_1)=\frac{\pi_1 p}{\pi_1 p + \pi_0 (1-p)}=\frac{1}{1+\frac{\pi_0(1-p)}{\pi_1 p}}
\end{equation}
and
\begin{equation}
P(s_1|q_0)=\frac{p (1-\pi_1)}{p(1-\pi_1) +  (1-p)(1-\pi_0)} = \frac{1}{1+ \frac{ (1-p)(1-\pi_0)}{p (1-\pi_1)}} 
\end{equation}
while $P(S=0|Q=1)$ by normalization. Accordingly:
\begin{multline}
\sigma_Q(p,\pi_0,\pi_1)=\\
h \circ f[\gamma/l(p)] \cdot t +
h \circ f[\gamma'/l(p)] \cdot (1-t)\,
%\theta(\frac{\pi_0(1-p)}{\pi_1 p}) [\pi_1 p + \pi_0 (1-p)]\\
%+ \theta(\frac{ (1-p)(1-\pi_0)}{p (1-\pi_1)}) [1-\pi_1 p + \pi_0 (1-p)]
\end{multline}
where $\gamma = \frac{\pi_0}{\pi_1}$, $\gamma' = \frac{1-\pi_0}{1-\pi_1}$, and $t = \pi_1 p + \pi_0 (1-p)$.

{\color{blue}{In other words:
$\sigma_Q(p,\pi_0,\pi_1):= G(\frac{1-p}{p} \frac{1-\pi_0}{1-\pi_1}) (1-p) + 
G(\frac{1-p}{p} \frac{\pi_0}{\pi_1}) p$}}

An alternative score might be achieved by replacing $h$ with $m$, i.e.,
\begin{equation}
\sigma_Q' := 
m \circ f[\gamma/l(p)] \cdot t +
m \circ f[\gamma'/l(p)] \cdot (1-t)\,
\end{equation}
As $f$ is monotone, we can commute $m$ and $f$ *** explain ***, i.e.,
\begin{equation}
\sigma_Q' := 
f \circ m[\gamma/l(p)] \cdot t +
f \circ m[\gamma'/l(p)] \cdot (1-t)\,
\end{equation}
with, as $\gamma>0$. $f \circ m[\gamma/l(p)] = f[\gamma m(l(p))]$
We can similarly define another score $\sigma_Q''$ based on $\tilde{i}_\epsilon$.
\subsection*{Explicit maximization}
$$H[S|Q]\simeq \sigma_Q' = \sum_q P(s^*(q)|q) P(q)$$
where $s^*(q):=\arg\max_s P(s|q)$. 
$$\sigma_Q'= \sum_q P(s^*(q),q) = \sum_q P(s^*(q) P(q|s^*(q))$$
Note that, if $s^*(q)$ is the same for each $q$, the score just corresponds to $P(s^*)$. In other words, the score is flat is each answer gives the same most likely skill level *** reasonable ***. Let us move to an imprecise setting and intend the $s^*$ state as
$s^*(q):=\arg\max_s \overline{P}(s|q)$. The objective function is a linear combination of $P(s*(q)$ whose coefficient can be optimized independently. Overall, this corresponds to a linear programming task.
$$\max_s \overline{P}(s,q) = \max_s \max_P P(s) P(q|s) = \max_s \overline{P}(s) \overline{P}(q|s)$$

\subsection*{Francesca's comments}

Comment1: This sentence is not clear to me:  "as $\gamma>0$. $f \circ m[\gamma/l(p)] = f[\gamma m(l(p))]$".\\
Comment2: How shall we use the Gumbel's trick

Comment 3: Below I obtain something that looks different to me, but maybe it is equivalent.

\begin{equation}
P(s_1|q_1)=\frac{\pi_1 p}{P(q_1)}
\end{equation}

\begin{multline}
\sigma_Q(p,\pi_0,\pi_1)=\\
m\left(\frac{\pi_1 p}{P(q_1)}\right) \cdot P(q_1)  +
m\left(\frac{(1-\pi_1) p}{1-P(q_1)}\right) \cdot (1-P(q_1))\,
\end{multline}

But, as $P(q_1)\geq 0$, then
\begin{multline}
m\left(\frac{\pi_1 p}{P(q_1)}\right) =\\
2max\left(\frac{\pi_1 p}{P(q_1)},1-\frac{\pi_1 p}{P(q_1)} \right) =\\
2max\left(\frac{\pi_1 p}{P(q_1)},\frac{\pi_0(1- p)}{P(q_1)} \right) =\\
\frac{2}{P(q_1)}max\left(\pi_1 p,\pi_0(1- p)\right).\\
\end{multline}

Similarly, as $1-P(q_1)\geq 0$, 
\begin{multline}
m\left(\frac{(1-\pi_1) p}{1-P(q_1)}\right) =\\
\frac{2}{1-P(q_1)}max\left((1-\pi_1) p,(1-\pi_0)(1- p)\right),\\
\end{multline}

so that,

\begin{multline}
\sigma_Q(p,\pi_0,\pi_1)= \\
2 max\{\pi_1 p, \pi_0 (1-p)\}+
2 max\{(1-\pi_1 )p, (1-\pi_0) (1-p)\}\,
\end{multline}

Comment 4: The optimisation in the credal setting shouldn't be the following one?

$$\max_P [\max_s P(s,q=0)+\max_s P(s,q=1)]$$

 $\P(s,q=0)$ and $P(s,q=1)$ are both functions of $p, \pi_0,$ and $\pi_1$ so I do not think we can split the optimization

$$\max_P \max_s P(s,q=0)+\max_P\max_s P(s,q=1)$$

However, it could be a reasonable approximation. Then, we can switch the optimization over $P$ and $s$ because they are both max, and obtain: 
$$\max_s \overline{P}(s)\overline{P}(q=0|s)+\max_s  \overline{P}(s)\overline{P}(q=1|s)$$.

If instead we are looking for the minimum over P, we cannot switch the optimization over $P$ and $s$:
$$\min_P \max_s P(s,q) \neq \max_s \min_P P(s,q)$$.

However I think we can split the optimization into 4 quadratic optimizations.
Let's use the following parametrization:
\begin{multline}\label{eq:minimization}
\min_{p,\pi_0,\pi_1} \sigma_Q(p,\pi_0,\pi_1) = \\
2\min_{p,\pi_0,\pi_1} \left( \max\{\pi_1 p, \pi_0 (1-p)\} + \\
~~~ \max\{(1-\pi_1 )p, (1-\pi_0) (1-p)\} \right) \approx \\
2\left(\min_{p,\pi_0,\pi_1}  \max\{\pi_1 p, \pi_0 (1-p)\} + \\
~~~ \min_{p,\pi_0,\pi_1} \max\{(1-\pi_1 )p, (1-\pi_0) (1-p)\} \right) 
\end{multline}

Let's consider the first optimization $\min_{p,\pi_0,\pi_1}  \max\{\pi_1 p, \pi_0 (1-p)\}$. It can be written as
$$ \min(\tilde{P}(s=0,q),\tilde{P}(s=1,q)) $$ 
where 
\begin{multline}
\tilde{P}(s=0,q) = \min_{\pi_1 ,p} \pi_1 p \\
\begin{array}{ll}
\text{subject to} & \pi_1 p > \pi_0 (1-p) \\
~&\underline{x} \leq x\leq \overline{x}\\ 
\end{array}
\end{multline}
and 
\begin{multline}
\tilde{P}(s=1,q) = \min_{\pi_0 ,p} \pi_0 (1-p) \\
\begin{array}{ll}
\text{subject to} & \pi_1 p < \pi_0 (1-p)
~&\underline{x} \leq x\leq \overline{x}\\ 
\end{array}
\end{multline}

where the constraint $x = (\pi_0, \pi_1 ,p)$.

In general, let $N_s$ be the number of skill levels and $N_q$ the number of possible answers, then the number of quadratic optimizations to solve should be equal to $N_s\cdot N_q$.

In this two dimensional example, we might not need the approximation above as the minimization in \ref{eq:minimization} can be written as $\min_{i=1:4}(\tilde{P}_1)$, where

\begin{multline}
\tilde{P}_1 = \min_{\pi_0,\pi_1,p} \pi_1 p+ (1-\pi_1)p = p \\
\begin{array}{ll}
\text{subject to} & \pi_1 p > \pi_0 (1-p) \\
~&  (1-\pi_1)p > (1-\pi_0)(1-p)  \\
~&\underline{x} \leq x\leq \overline{x}\\ 
\end{array}
\end{multline}
and so on. However, in this case, the number of optimizations to solve should be $N_s^{N_q}$, that is, it grows exponentially with  $N_q$.

\paragraph{*** Alessandro's reformulation ***}
\begin{equation}
\begin{split}
\text{min} \displaystyle \, \pi_1 \cdot p \\
\text{subject to} & \displaystyle \pi_1 p > \pi_0 (1-p)\\
                 &\displaystyle \underline{\pi}_0 \leq \pi_0 \leq \overline{\pi}_0 \\
                 &\displaystyle \underline{\pi}_1 \leq \pi_1 \leq \overline{\pi}_1 \\
                 &\displaystyle \underline{p} \leq p \leq \overline{p} \\
\end{split}
\end{equation}
with the new variable specification:
\begin{eqnarray}
x_1&:=&\pi_1 \cdot p \,,\\
x_2&:=&\pi_0 \cdot p \,,\\
x_3&:=&(1-\pi_1) \cdot p \,,\\
x_4&:=& \pi_0 \cdot (1-p)\,.\\
\end{eqnarray}
the problem becomes:
\begin{equation}
\begin{split}
\text{min} \displaystyle \, x_1 \\
\text{subject to\phantom{xx}} & \displaystyle x_1 > x_4\\
                 &\displaystyle \underline{\pi}_0 \leq x_2+x_3 \leq \overline{\pi}_0 \\
                 &\displaystyle \underline{\pi}_1 \leq x_1+x_4 \leq \overline{\pi}_1 \\
                 &\displaystyle \underline{p} \leq x_1+x_3 \leq \overline{p} \\
\end{split}
\end{equation}

\section{Introduction}\label{sec:intro}
The use of communication and information technologies in education is actually growing. Both online (e.g., MOOCs) and classroom courses are urgently asking for more flexible and sophisticated e-learning and e-testing tools \cite{pollard2001exploring}. AI-based approaches such as intelligent tutoring systems, adapting the interaction with the student on the basis of his/her knowledge and/or psychological profile, represent an important direction to improve the quality of the (e-)learning experience \cite{burns2014intelligent}.

Bayesian networks (BNs) \cite{koller2009} have been used to model the knowledge driving such intelligent systems \cite{almond2015bayesian}. However, collecting large sets of reliable data in educational domains may be difficult and time consuming (e.g., a course with few students, or taught for the first time), and the quantification should be based on expert knowledge only. To elicit a Bayesian network, an expert might face questions like: ``\emph{which is the probability of a student with a particular knowledge level giving the right answer to a question?}''. Giving sharp probabilities for questions of this kind can be problematic for an expert, whose knowledge is mostly qualitative (e.g., ``\emph{a right answer is very unlikely}''). Fuzzy linguistic approaches represent a viable, non-numerical, way to address these issues \cite{badaracco2013fuzzy}. To stick within the probabilistic framework, verbal-numerical probability scales associated with sharp values \cite{renooij1999talking} or intervals \cite{walley1991statistical} have been also proposed.

In this paper we show how to conjugate an interval-valued probabilistic elicitation of expert knowledge with the BN framework. This means to cope with a \emph{credal network} (CN) \cite{novapiatti}, a generalization of BNs based on the imprecise probability theory \cite{walley1991statistical}, where local parameters are defined by set-valued probabilities. This simplifies the elicitation process and offers a more reliable handling of the related uncertainty. Moving from BNs to CNs implies two main issues: (i) numerical inferences will be interval-valued too, thus making debatable both the decision criterion \cite{troffaes} an the information measures \cite{klir1999uncertainty} to adopt; and (ii) inference tasks in CNs typically belongs to higher complexity classes than their Bayesian counterparts \cite{maua14jair}. Both these issues are addressed by defining a computationally feasible procedure based on CNs to be used for practical implementation of intelligent systems solely specified by expert knowledge. To the best of our knowledge this is the first attempt to perform e-testing with models of this kind.

We focus on the application of CNs to \emph{computer adaptive testing} (CAT), i.e., an approach to e-testing that adjusts the sequence and the number of questions to the ability level of the test taker. CATs have the potential to make the test an individualised experience that challenges and does not discourage the test takers, as most of the questions are near their ability levels. Building upon \emph{item response theory} \cite{hambleton1985item}, the common background underpinning CATs, graphical modeling (such as BNs and CNs) offers a powerful language for describing complex multivariate dependencies between skills and rich tasks. Several researchers have exploited the potential of BNs both in adaptive and non-adaptive educational assessment \cite{vomlel2004building,plajner2015}. These authors focus on applications for which data are available to learn the model parameters. We regard this point as a serious limitation, possibly hindering CATs adoption by many teachers and instructors.

We start from a CAT procedure based on BNs that uses \emph{entropy} as the information-theoretic measure driving the question selection and the stopping criteria (Sect.~\ref{sec:bayesianCAT}). Our goal is to improve  this procedure by using CNs to better describe the pervasive uncertainty characterizing the model. A direct extension of the Bayesian framework to CNs would require the computation of bounds for the conditional entropy with respect to the CN specification. This corresponds to a non-linear non-convex optimization task. We therefore propose a number of simplifying assumptions to overcome this problem at the price of accepting sub-optimal question selection schemes (Sect.~\ref{sec:credalCAT}). The approach is tested on a real-world benchmark about German language proficiency assessment (Sect.~\ref{sec:conc}). The results are promising: CAT based on CNs is effective in reducing the number of questions while maintaining a high accuracy in the evaluation and the approximations introduced do not compromise the procedure's effectiveness.

\section{Bayesian CAT}

\paragraph{Skills modeling.} We describe the knowledge level of a student as a collection of categorical variables, say $\bm{X}:=(X_1,\ldots,X_n)$, called \emph{skills}. A joint \emph{probability mass function} (PMF) $P(\bm{X})$ describes the uncertainty about the actual values of the skills. A compact specification of such multivariate model can be achieved by a BN \cite{koller2009}. This corresponds to: (i) a directed acyclic graph whose nodes are in one-to-one correspondence with the variables of $\bm{X}$; and (ii), for each $X_i\in\bm{X}$, 
 a collection of conditional PMFs $P(X_i|\mathrm{\pi}_{X_i})$, one for each value $\pi_{X_i}$ of the joint variable $\mathrm{\Pi}_{X_i}$ denoting the \emph{parents} (i.e., the immediate predecessors) of $X_i$. The \emph{Markov condition} for BNs assumes every variable conditionally independent of its non-descendants non-parents given the parents. Accordingly, the joint PMF associated with a BN is such that $P(\bm{x}):= \prod_{i=1}^n P(x_i|\pi_{X_i})$, for each $\bm{x}$, where the values of $x_i$ and $\pi_{X_i}$ are those consistent with $\bm{x}$.

\paragraph{Questions modeling.} The above joint probabilistic model describes the uncertainty about the skills of a student prior to his/her answers to the questions. To evaluate the student we formulate a number of \emph{questions}, described as a collection of variables $\bm{Y}:=(Y_1,\ldots,Y_m)$. We assume these variables to be Boolean, with the true value corresponding to the correct answer.\footnote{Extension to non Boolean answers is trivial as all answers $Y_i$ are \emph{manifest} variables, and, thus, $Y_i$ can be always regarded as a binary variable with the two values denoting the observed answer $y_i$ and its negation \cite{antonucci2009}.} We call \emph{background} of a question the set of skills ``required'' to answer it. This can be regarded as a conditional independence statement: given the background skills, the answer to the question is independent of the other skills and of the other questions. Following the Markov condition, this can be modeled by representing each question as a leaf node whose parents are the background skills. Such augmented graph requires the quantification, for each $Y_j\in\bm{Y}$, of a conditional PMF $P(Y_j|\mathrm{\pi}_{Y_j})$ for each value $\mathrm{\pi}_{Y_j}$ of the background skills $\mathrm{\Pi}_{Y_j}$. This procedure defines a BN over the skills and the questions, and hence a joint PMF $P(\bm{X},\bm{Y})$.

\paragraph{Non-adaptive testing.} Let $\bm{Y}=\bm{y}$ denote a student's answers to the test. In the above considered framework, the posterior knowledge about the skills is modeled by the joint PMF $P(\bm{X}|\bm{y})$. By running standard BN updating algorithms, the most probable level $\tilde{x}_i$ of skill $X_i$ can be therefore evaluated  as $\tilde{x}_i:= \arg\max_{x_i} P(x_i|\bm{y})$, for each $X_i\in\bm{X}$. This reflects a non-adaptive, probabilistic approach to student evaluation.

\paragraph{Adaptive testing.} To add adaptiveness to the above approach, every question should be chosen on the basis of the previous answers. As the goal is to gather information about the student skills, we evaluate the expected \emph{information gain} (IG, i.e., the change in information entropy) associated with each possible new question, and pick the one maximizing this measure. The entropy of a BN over $\bm{X}$ can be computed as $H(\bm{X}):=\sum_{i=1}^n H(X_i|\mathrm{\Pi_{X_i}})$ \cite{koller2009}, where $H(X_i|\mathrm{\Pi_{X_i}}):=\sum_{\pi_{X_i}} H(X_i|\pi_{X_i}) P(\pi_{X_i})$ is the \emph{conditional entropy} for $X$ given its parents and $H(X_i|\pi_{X_i})$ is the entropy of the conditional PMF $P(X_i|\pi_{X_i})$.\footnote{To have entropy levels between zero and one, we define the entropy of the PMF $P(X)$ as $H(X):=-\sum_x P(x) \log_b P(x)$, with $b$ number of states of $X$.} Let $\bm{Y}=\bm{y}$ denote the answers to the questions already asked and $\bm{Y}'$ the set from which the next question should be picked. If the answer to every question $Y'\in\bm{Y}'$ would be known, and denoted by $y'$, the question $\tilde{Y}' \in \bf{\mathbf{Y}'}$ to choose would be the one leading to the largest IG. Yet, as the decision has to be made before the student's answer, conditional entropy should be considered instead, i.e.,
\begin{equation}
\tilde{Y}' := \arg \max_{Y'\in \bm{Y}'} \left[H(\bm{X}|\bm{y})-H(\bm{X}|Y',\bm{y})\right]\,. 
\label{eq:expent}
\end{equation}

\begin{figure*}[htp!]
\centering
\begin{tikzpicture}[auto,scale=.7]
\node[block,node distance=.cm,fill=black!10] (pick) {\begin{varwidth}{1.5cm}\tiny\centering PICK A QUESTION \end{varwidth}};
\node[cloud, left of=pick,node distance=3cm,fill=black!5] (db) {\begin{varwidth}{1.5cm}\tiny\centering QUESTIONS DATABASE\end{varwidth}};
\node[block, right of=pick,node distance=3cm,fill=black!10] (ask) {\begin{varwidth}{1.5cm}\tiny\centering ASK QUESTION\end{varwidth}};
\node[block, right of=ask,node distance=3cm,fill=black!10] (collect) {\begin{varwidth}{1.5cm}\tiny\centering COLLECT ANSWER\end{varwidth}};
\node[decision, below of=collect,node distance=1.4cm] (decide) {\begin{varwidth}{1.5cm}\tiny\centering \textbf{STOPPING RULE? }\end{varwidth}};
\node[block, left of=decide, node distance=3cm,fill=black!10] (stop) {\begin{varwidth}{1.5cm}\tiny\centering END TEST\end{varwidth}};
\node[block, left of=stop, node distance=3cm,fill=black!10] (assign) {\begin{varwidth}{1.5cm}\tiny\centering ASSIGN LEVELS \end{varwidth}};
\path[line] (decide) -- node {\tiny YES}(stop);
\path[line] (decide) -- node[pos=0.1,above] {\tiny NO}(pick);
\path[line,dashed] (db) -- (pick);
\path[line] (pick) -- (ask);
\path[line] (ask) -- (collect);
\path[line] (collect) -- (decide);
\path[line] (decide) -- (stop);
\path[line] (stop) -- (assign);
\end{tikzpicture}
\caption{CAT procedure}
\label{fig:alg}
%\vspace{-1cm}
\end{figure*}

CATs should also decide when to stop asking questions. Again, entropy can be used as a measure to decide when the current evaluation is sufficiently informative, i.e., we stop the test if the skills entropy given the answers is below some threshold $\tilde{H}$. The overall approach is depicted in Fig.~\ref{fig:alg}.

\section{Credal CAT}

\paragraph{Credal sets and credal networks.} A set of PMFs over $X_i$ is called here \emph{credal set} (CS) and denoted as $K(X_i)$. We always remove the inner points (i.e., those corresponding to convex combinations of the others) of a CS. CNs \cite{cozman2000} are generalized BNs whose local PMFs are replaced by CSs.  The BN defined in the previous section over the skills $\bm{X}$ and the questions $\bm{Y}$ becomes a CN if we replace with CSs the skill-to-skill and skill-to-question conditional PMFs. A joint CS $K(\bm{X},\bm{Y})$ is consequently obtained as the collection of all the joint PMFs induced by BNs whose parameters take their values from the corresponding CSs, i.e., 
\begin{equation}\label{eq:strongext}
K(\bm{X},\bm{Y}):=\left\{ P(\bm{X},\bm{Y}) \left| 
\substack{
P(\bm{x},\bm{y}):=\prod_{i=1}^n P(x_i|\pi_{X_i}) \cdot \prod_{j=1}^m P(y_j|\pi_{Y_j}), \\P(X_i|\pi_{X_i}) \in K(X_i|\pi_{X_i}) ,\, P(Y_j|\pi_{Y_j}) \in K(Y_j|\pi_{Y_j})}
\right.
\right\}\,,
\end{equation}
where the values of $x_i$, $\pi_{X_i}$, $y_j$, $\pi_{Y_j}$ are those consistent with $\bm{x}$ and $\bm{y}$.

\paragraph{Expert knowledge modeling.} For a reliable expert knowledge modeling, we use CSs induced by \emph{probability intervals}. Qualitative judgments about the probability of a state are converted in interval constraints such as $l\leq P(x) \leq u$, with the interval $[l,u]$ capturing the expert knowledge behind the judgment in a more reliable way than a sharp assessment. The CS consistent with these constraints is eventually obtained by standard polyhedral algorithms. Verbal to interval-numeric scales such as that Tab.~\ref{tb:labels} are used. For instance, if for the probability of the true state of the Boolean variable $Y$ the expert judgment is ``very likely'', the corresponding linear constraint is $.2 \leq P(Y=\textrm{true}) \leq .4$.% and we obtain a CS $K(Y)$ with the PMFs $P_1(Y)=[.2,.8]$ and $P_2(Y)=[.4,.6]$ as non-inner points.

\paragraph{Non-adaptive testing.} Given the answers $\bm{y}$ to the questions $\bm{Y}$, we evaluate the student as in the previous section by updating the marginal probabilities of each skill. With CNs, these posterior values are set-valued and their characterization can be provided by lower and upper bounds, say $\underline{P}(X_i|\bm{y})$ and $\overline{P}(X_i|\bm{y})$ for each $X_i\in\bm{X}$. CN updating algorithms can eventually compute these bounds. The task displays higher complexity than in the case of BNs (e.g., exact inference in non-binary singly-connected CNs is NP-hard \cite{maua14jair}), but approximate techniques can be considered when exact inference is unfeasible \cite{antonucci2014e}.

To compare the posterior intervals and decide the actual level of the student we might adopt the (conservative) \emph{interval dominance} criterion \cite{troffaes}, which rejects a level if its upper probability is smaller than the lower probability of some other level. Overlaps between intervals might therefore induce a situation of \emph{indecision} between two or more levels. This is a so-called \emph{credal} classification of the student level \cite{zaffalon2012evaluating}, and it represents the fact that students answers are somehow contradictory or not informative enough to provide a sharp decision. Interval dominance can return unnecessarily imprecise results. \emph{Maximality} is a more refined criterion that rejects the levels which are less probable than another level for all the elements of the CS \cite{walley1991statistical}. Maximality  can be reduced to multiple updating tasks on auxiliary binary leaf nodes defined for each pair of states \cite{antonucci2014e}.

\paragraph{Adaptive testing.}
To achieve CAT with CNs using entropy as measure of informativeness for PMFs, as in the BN approach of Sect. \ref{sec:bayesianCAT}, computation of entropies should be extended to CSs. This topic has been the subject of much discussion \cite{klir1999uncertainty}. A cautious approach \cite{abellan2003maximum} consists in taking the upper entropy $\overline{H}(\bm{X})$, i.e., the entropy of the most entropic PMF in the convex closure $\overline{K}(\bm{X})$ of  $K(\bm{X})$. 
In our framework, we should, then, look for maximum values of conditional entropies, such as $\overline{H}(X_i|Y',\bm{y})$ or $\overline{H}(X_i|\Pi_{X_i})$, as conditional entropies are required to compute both: (i) the joint (unconditional) entropy $H(\bm{X})$ (and its posterior values); and (ii) the conditional entropies  involved in the  question selection in Eq.~\eqref{eq:expent}.
By definition a conditional entropy is a convex combination (whose weights are the elements of a marginal PMF) of convex functions (the entropies). The objective function might, then, be non-convex, as the weights are also optimization variables.\footnote{E.g., if $f(x)$ and $g(x)$ are convex functions of $x$, $h(x,y):= y f(x) + (1-y) g(x)$ is not convex even for $0 \leq y \leq 1$.}

Then, to bypass this non-convex optimization task, we compute (i) by separately considering the entropies of each skill $X_i\in\bm{X}$. This is analogous to the marginal approach commonly considered in multi-label classification to minimize Hamming losses \cite{antonucci2016c}. The issue (ii) is more challenging. We consider the following upper approximation of $\overline{H}(X_i|Y',\bm{y})$:
\begin{equation}
\overline{\overline{H}}(X_i|Y',\bm{y})=
\max_{P(y'|\bm{y})\in
\{\underline{P}(y'|\bm{y}),\overline{P}(y'|\bm{y})\}}
\sum_{y' \in \{\textrm{true},\textrm{false}\}}
\overline{H}(X_i|\bm{y},y') P(y'|\bm{y})\,,
\end{equation}
where the bounds of $P(y'|\bm{y})$ are obtained by standard CN updating algorithms.
The problem thus reduces to the computation of upper entropies as
\begin{equation}\label{eq:entropy}
\overline{H}(X_i|\bm{y}) := \sup_{P(X_i|\bm{y}) \in \overline{K}(X_i|\bm{y})} H(X_i|\bm{y})\,,
\end{equation}
where  $\overline{K}(X_i|\bm{y})$ is the posterior CS after conditioning on the observed answers $\bm{y}$.  If $\overline{K}(X_i|\bm{y})$ has a finite number of non-inner points, this is a linearly-constrained convex optimization whose solution typically corresponds to either the uniform PMF or a non-inner point on the frontier of $\overline{K}(X_i|\bm{y})$. A numerical solution can be easily found by a simple iterative approach in the special case of CS specified by probability intervals \cite{abellan2003maximum}. We have therefore computed the posterior lower and upper bounds of $P(X_i|\bm{y})$, and then maximized the entropy with respect to those bounds. % as in Eq.~\eqref{eq:entropy}. 
The procedure induces an outer approximation of $\overline{K}(X_i|\bm{y})$, and hence the upper approximation  of the maximum entropy $\overline{\overline{H}}(X|\bm{y})\geq\overline{H}(X|\bm{y})$. 
%With a small notation abuse such approximate value is also denoted as $\overline{H}(X|\bm{y})$ and is regarded here as a measure of the informativeness level achieved for skill $X$ after the answers $\bm{y}$. 
Finally, to generalise Eq.~\eqref{eq:strongext} to CNs, we define the information gain provided by a question $Y'$ for its background skill $X_{Y'}$ as $\overline{\overline{H}}(X_{Y'}|\bm{y})-\overline{\overline{H}}(X_{Y'}|Y',\bm{y})$ and select the question $\tilde{Y}'$ leading to the maximum information gain, i.e.,
\begin{equation}\label{eq:credalentropy}
\tilde{Y}' := \arg \max_{Y'\in\bm{Y}'} 
\left[
 \overline{\overline{H}}(X_{Y'}|\bm{y})-\overline{\overline{H}}(X_{Y'}|Y',\bm{y}) 
 \right]
 \,.
\end{equation}
For the stopping criterion, as we do not consider the joint entropy over the skills, we separately require each $\overline{\overline{H}}(X_i|\bm{y})$ to be smaller than a threshold $\tilde{H}$. To be consistent with this choice, we remove from the set of questions to be selected, those whose background skills already satisfy this condition.

Note that the use of an outer approximation of the upper entropy affects only the question selection process (eventually making it sub-optimal), whereas it has no effect on the student evaluation given a set of answers.

\section{Application}
Before the academic year begins, the students of the University of Applied Sciences and Arts of Southern Switzerland (SUPSI) are asked to take an online German language placement test with 95 questions. In years 2015 and 2016, the answers of 451 students to all the questions have been collected. This benchmark is used to simulate CATs based on BNs and CNs as described in Sects.~\ref{sec:bayesianCAT} and \ref{sec:credalCAT}.

\paragraph{Model elicitation.}
Four skills are assessed: \textit{W\"ortschatz} ($X_1$, vocabulary), \textit{Kommunikation} ($X_2$, communication), \textit{H\"oren} ($X_3$, listening), and \textit{Lesen} ($X_4$, reading). For each skill the student is assigned to a knowledge level compliant with EU guidelines.\footnote{\url{http://www.coe.int/t/dg4/linguistic/Source/Framework_EN.pdf}.}
Levels A1, A2, B1, and B2 are considered, and skills are therefore modeled as quaternary variables.
Teachers associate each question with a single skill, which is set as the unique background skill of the question. The number of questions associated with $X_1$/$X_2$/$X_3$/$X_4$ is 26/24/30/15. The current evaluation method assigns levels by setting thresholds on the percentage $\gamma$ of correct answers on each skill (A1 if  $\gamma < 35\%$, A2 up to $55\%$, B1 up to $75\%$).\footnote{These data as well as the software used for the simulations are freely available at \url{http://ipg.idsia.ch/software.php?id=138}.}

We first elicit from the teachers the structure of the BN/CN graph over the skills. The result is a chain, which is augmented by leaf nodes modeling the questions, each having its background skill as single parent. Overall, a tree-shaped topology as in Fig.~\ref{fig:fourskills} is obtained. This makes exact inference in the BN fast, while in the CN a variable elimination might be slow (a minute for query in our setup). A faster approximate CN algorithm is therefore used \cite{antonucci2014e}.

\begin{figure}[htp!]
\centering
\begin{tikzpicture}[scale=.8]
\node[nodo] (w)  at (-0.25,0) {$X_1$};
\node[nodo] (k)  at (3,0) {$X_2$};
\node[nodo] (h)  at (6.5,0) {$X_3$};
\node[nodo] (l)  at (9.25,0)  {$X_4$};
\draw[arco] (w) -- (k);
\draw[arco] (k) -- (h);
\draw[arco] (h) -- (l);
\node[Qnodo] (f1)  at (-0.25,-1) {$Y_1$};
\node[Qnodo] (f2)  at (1.5,-1) {$Y_2$};
\node[Qnodo] (f3)  at (3,-1) {$Y_3$};
\node[Qnodo] (f4)  at (4.5,-1) {$Y_4$};
\node[Qnodo] (f5)  at (6.5,-1) {$Y_5$};
\node[Qnodo] (f6)  at (8.5,-1)  {$Y_6$};
\node[Qnodo] (f7)  at (10,-1)  {$Y_7$};
\draw[arco] (w) -- (f1);
\draw[arco] (k) -- (f2);
\draw[arco] (k) -- (f3);
\draw[arco] (k) -- (f4);
\draw[arco] (h) -- (f5);
\draw[arco] (l) -- (f6);
\draw[arco] (l) -- (f7);
\end{tikzpicture}
\caption{A directed graph for CAT\label{fig:fourskills}.}
\end{figure}
\vspace{-5mm}

Teachers report their knowledge about the unconditional states of $X_1$ and the conditional states of $X_i$ given $X_{i-1}$, for $i=2,3,4$, as qualitative judgments (top of Tab.~\ref{tb:CPTs}). 
To simplify the elicitation, the probabilities $P(X_i|X_{i-1})$ are given the same verbal judgment for all $i=2,3,4$. A more detailed model could provide more accurate evaluations but it would be very hard for the domain expert to elicit it in a reliable way. 
Also, questions are divided by the teachers in three groups, corresponding to different difficulty levels. Questions in the same group are quantified in the same way, irrespective of their background skill, giving the judgments reported in the bottom part of Tab.~\ref{tb:CPTs}.

{\small \begin{table*}[htp!] 
\renewcommand{\arraystretch}{1.2}
\centering
{%\footnotesize
\begin{tabular}{@{}lc@{}}
\toprule
$X_1$&$P(X_1)$\\
\midrule
A1&\emph{improbable}\\
A2&\emph{uncertain}\\
B1&\emph{uncertain}\\
B2&\emph{improbable}\\
\bottomrule
\end{tabular}
\quad
\begin{tabular}{@{}lcccc@{}}
\toprule	
$P(X_i|X_{i-1})$&\phantom{.}$X_{i-1}=\mathrm{A1}$\phantom{.}&\phantom{.}$X_{i-1}=\mathrm{A2}$\phantom{.}&\phantom{.}$X_{i-1}=\mathrm{B1}$\phantom{.}&\phantom{.}$X_{i-1}=\mathrm{B2}$\phantom{.}\\
\midrule
$X_i$=A1&\emph{fifty-fifty}&\emph{uncertain}&\emph{improbable}&\emph{impossible}\\
$X_i$=A2&\emph{uncertain}&\emph{fifty-fifty}&\emph{uncertain}&\emph{improbable}\\
$X_i$=B1&\emph{improbable}&\emph{uncertain}&\emph{fifty-fifty}&\emph{uncertain}\\
$X_i$=B2&\emph{impossible}&\emph{improbable}&\emph{uncertain}&\emph{fifty-fifty}\\
\bottomrule
\end{tabular}}
\vskip 2mm
\begin{tabular}{@{}lcccc@{}}\toprule
$P(Y=T|X)$\phantom{aa}&\phantom{a}$X=\mathrm{A1}$\phantom{a}&\phantom{a}$X=\mathrm{A2}$\phantom{a}&\phantom{a}$X=\mathrm{B1}$\phantom{a}&\phantom{a}$X=\mathrm{B2}$\phantom{a}\\
\midrule
Easy&\emph{uncertain}&\emph{fifty-fifty}&\emph{expected}&\emph{probable}\\
Medium&\emph{improbable}&\emph{uncertain}&\emph{fifty-fifty}& \emph{expected}\\
Difficult&\emph{impossible}&\emph{improbable}&\emph{uncertain}& \emph{fifty-fifty}\\
\bottomrule
\end{tabular}
\vskip 2mm
\caption{Expert judgements.}\label{tb:CPTs}
\end{table*}}
%\vspace{-10mm}
For the CN, those judgements are translated in interval constraints for the corresponding events on the basis of the verbal-numerical scale in Tab.~\ref{tb:labels}. Different probability intervals are considered for skills and questions as they refer to events of different type. For instance, when the expert considers ``\emph{impossible}'' for an A1 level student to know the answer to a difficult question, the student is assigned a probability between .175 and .2 of answering correctly, as the questions offer only four choices plus the option of giving no answer. 
Notice that, by doing so, we are not anymore assuming that all questions in the same difficulty group share exactly the same conditional PMFs (as done by the BN model), as PMFs of different questions can vary independently in the given intervals. This seems a more sensible assumption than that of the precise model.
For the BN, the PMFs corresponding to the centers of mass of the CSs defining the CN are used. Numerical inferences in the BN are consequently included in the intervals computed with the CN.

{\small
\begin{table*}[htp!] 
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{@{}lcccccc@{}}
\toprule
Judgement&\emph{impossible}&\emph{improbable}&\emph{uncertain}&\emph{fifty-fifty}&\emph{expected}&\emph{probable}\\
\midrule
Skills&$1$-$10\%$&$10$-$20\%$&$20$-$40\%$&$30$-$50\%$&-&-\\
Questions&$17.5$-$20\%$&$22.5$-$25\%$&$30$-$35\%$&$60$-$65\%$&$75$-$80\%$& $95$-$97.5\%$\\
\bottomrule
\end{tabular}
\caption{A verbal-numerical scale for probability-intervals elicitation.}\label{tb:labels}
\end{table*}}

\paragraph{Experimental results.} 
%Acronyms BN-NA, CN-NA, BN-AD, CN-AD denote 
BN and CN methods in their non-adaptive (NA) and adaptive (AD) versions are considered. \emph{Accuracy}, i.e., the proportion of students to whom the test assigns the same level of the current evaluation method, describes BN performances. This measure cannot be used for the set-valued outputs of CN methods. In this case the $u_{65}$ measure can provide a comparison with the accuracy \cite{zaffalon2012evaluating}. If $\mathcal{L}$ is the set of levels assigned by the CN on a skill and $L$ its cardinality, a \emph{discounted} accuracy gives $1/L$ if $\mathcal{L}$ includes the true level and zero otherwise. The $u_{65}$ is a concave reinforcement of this score based on risk-adverse arguments. Its underlying assumption is that acknowledging the indecision between more levels has larger utility than randomly choosing one of them (e.g., the teacher could set up further assessments in the undecided cases). Tab.~\ref{tb:nonAdaptiveResults} shows the NA comparison. In Fig.~\ref{fig:plots} (left), the BN-NA accuracy is separately evaluated on the determinate (light bars) and indeterminate (dark bars) instances, i.e. those for which, respectively, a single level or multiple levels are returned by the CN model. On average, CN-NA returns single levels in $37.25\%$ of the cases and, if this is not the case, an average of $2.36$ levels ($3.22$ with interval dominance) are returned.

\begin{table*}[htp!]
	\centering
	\begin{tabular}{@{}lccccc@{}}\toprule
		Algorithm&\phantom{aa}Average\phantom{aa}&\phantom{aaa}$X_1$\phantom{aaa}&\phantom{aaa}$X_2$\phantom{aaa}& \phantom{aaa}$X_3$\phantom{aaa}&\phantom{aaa}$X_4$\phantom{aaa}\\
		\midrule
		BN-NA (acc)&$63.09\%$&$	67.56	\%$&$60.85\%$&${	75.84	\%}$&$48.10\%$\\
		CN-NA ($u_{65}$)& ${65.37\%}$&$67.71\%$&${66.67\%}$&$	70.33\%$&${56.76\%}$\\
		\bottomrule
	\end{tabular}
	\caption{Non-adaptive tests results.}\label{tb:nonAdaptiveResults}
\end{table*}

In the AD case we also track the average number of asked questions. Results are in Fig.~\ref{fig:plots} (right). CN-AD (circles) is tested for different thresholds over the entropy (labels of the markers) against a version of BN-AD based on the joint entropy (triangles). Similar values are obtained by coping with marginal entropies. We also allow the BN-AD method to return multiple levels by maximizing the expected $u_{65}$ utility over any possible set of levels. This variant is called BN-AD' and the corresponding $u_{65}$ measure is reported  (squares).

\begin{figure}[htp!]
\centering
\begin{tikzpicture}
\begin{axis}[width=4cm,height=3.5cm,ybar=0,bar width=7pt,ylabel={\small acc}, xlabel={Skills},xtick=data,ytick={0,25,50,75,100},ymin=0,ymax=100,yticklabels={0,.25,.50,.75,1},symbolic x coords={$X_1$,$X_2$,$X_3$,$X_4$}, y label style={at={(axis description cs:0.3,.5)},anchor=south},xtick align=inside, legend style={at={(1,1)},anchor=south east},legend style={draw=none}]
\addplot[fill=black!10] coordinates {($X_1$,85.42)($X_2$,87.06)($X_3$,89.51)($X_4$,56.34)}; 
\addlegendentry{\tiny{determinate}} 
\addplot[fill=black!70] coordinates {($X_1$,54.51)($X_2$,45.13)($X_3$,67.72)($X_4$,44.26)};
\addlegendentry{\tiny{indeterminate}}
\end{axis}
\end{tikzpicture}
\begin{tikzpicture}
\begin{axis}[width=7.5cm,height=4cm,ylabel={\small acc/$u_{65}$},xlabel={Number of questions},xmin=35, xmax=100,ymin=50, ymax=70, xtick={40,50,60,70,80,90}, ytick={50,55,60,65},legend style={at={(1,0)},anchor=south east},ymajorgrids=true,xmajorgrids=true,  y label style={at={(axis description cs:0.13,.5)},anchor=south},legend style={draw=none}]
\addplot[black,thick,mark=triangle,mark options={fill=black}, visualization depends on=\thisrow{alignment} \as \alignment, nodes near coords, point meta=explicit symbolic,  every node near coord/.style={anchor=\alignment}] 
table [meta index=2 ] {
x y label alignment
40.36	51.51 .7 0
49.84	55.43 .65 -20
58.34	57.49 .6 80
66.50 59.45 .55 80
79.43	62.75 .45 80
89.04	63.14 .3 80
95  63.09     0  160
};
\addlegendentry{\tiny{BN-AD}}
\addplot[magenta,thick,mark=square,mark options={fill=black}, visualization depends on=\thisrow{alignment} \as \alignment, nodes near coords, point meta=explicit symbolic,  every node near coord/.style={anchor=\alignment}] 
table [meta index=2 ] {
x y label alignment
40.36	59.9 .7 -20
49.84	61.00 .65 -40
58.34	60.86 .6 50
66.50 60.34 .55 -30
79.43	64.90 .45 -80
89.04	66.3 .3 -80
95  66.22     0  -160
};
\addlegendentry{\tiny{BN-AD'}}
\addplot[dashed,blue,thick,mark=o,mark options={solid,fill=white},
visualization depends on=\thisrow{alignment} \as \alignment,
nodes near coords, % Place nodes near each coordinate
point meta=explicit symbolic, % The meta data used in the nodes is not explicitly provided and not numeric
every node near coord/.style={anchor=\alignment} ] table [meta index=2 ] {
x       y   label  alignment
42.45 58.27  .95   20
56.96 62.37  .925   -80
66.87 63.49   .9   -80
83.75 62.83  .8    -110
95   65.37   0  160
};
\addlegendentry{\tiny{CN-AD}}
\end{axis}
\end{tikzpicture}
\caption{Non-adaptive (left) and adaptive (right) tests performance.}\label{fig:plots}
\end{figure}

As a comment, CNs seem to identify hard-to-evaluate students as those for which multiple levels are provided. In fact, the agreement between the BN and the traditional tests is larger when the CN test is determinate. As a consequence, the CN $u_{65}$ measure is, on average, larger than the BN accuracy. 
A limitation of the CN test is the large fraction of indeterminate evaluations. One can interpret this result as a lack of robustness of the BN model, as even small variations in the model specifications can result in different decisions. Results also show that, both BN-AD and CN-AD approaches reduce the number of questions asked without significantly affecting the accuracy. BN-AD performances are improved by the ``credal'' variant BN-AD'. The results becomes very similar to those of the CN-AD. Yet, the latter method appears to be a more principled and suitable approach for a direct modeling of qualitative expert knowledge.

%does not require any, potentially questionable, imputation of the probabilistic ranges provided by the experts.

%Yet, we regard theof BN-AD, we regard the CN approach as more principled and suitable for a direct modeling of qualitative expert knowledge.

%uch a difference disappear if the credal-like variant BN-AD' is considered.
%However, the CN-AD procedure appears to be more effective, as the $u_{65}$ accuracy measure decreases more slowly than the accuracies of BN-AD procedures.  
%Finally,comparing the two BN-AD approaches, one can see that the use of the local instead of the joint entropy does not compromise the effectiveness of the adaptive approach.

%Different decision criteria for the assignment of the skill level, e.g., maximality or e-admissibility, should be considered.
%As a consequence, although BNs achieve better performance when considering the discounted accuracy, CN tests should be preferred whenever the utility is better described by the U65 or U80 measures, that is, whenever being undecided between more level should be preferred than assigning one of them at random. 
%As the true level of the students is unknown, this measure should be regarded as the agreement between to different evaluation strategies. 
%\begin{mdframed}[hidealllines=true,backgroundcolor=blue!20]
%\paragraph{Experimental results.} 
%Tab.~\ref{tab:nonadaptive} shows in the non-adaptive case the relative agreement between the traditional evaluation method (TEM) and the two BN-based models (IBN and TBN), computed as the proportion of students to which the tests under comparison assign the same level, averaged over 10 runs of cross-validation. On average 15\% of the students are assigned to a different level with respect to the traditional method (although the difference is never more than one level). The reason is that the BN models account for the difficulty of the questions. E.g., a student with only 70\% of correct answers could be assigned the level B2 if he/she has answered correctly a large number of difficult questions which a student of lower level could not have guessed just by chance. 
%\end{mdframed}
%
%\begin{table}[htp!]
%%\renewcommand{\arraystretch}{1.1}
%\centering
%\begin{tabular}{@{}p{2.5cm}p{1.2cm}p{1.2cm}p{1.2cm}p{1.2cm}p{1.2cm}@{}}\toprule
%Algorithm&\multicolumn{1}{l}{W\"ort.}&\multicolumn{1}{l}{Komm.}&\multicolumn{1}{l}{H\"oren}&\multicolumn{1}{l}{Lesen}&\multicolumn{1}{l}{Overall}\\ 
%\midrule
%TEM vs IBN&.80 &.87 &.89 &.85 &.85 \\
%TEM vs TBN&.79 &.87 &.88 &.83 &.84 \\
%IBN vs TBN&.98 &.95 &.94 &.92 &.95 \\
%\bottomrule
%\end{tabular}
%\caption{Relative agreement between models}\label{tab:nonadaptive}
%\vspace{-0.cm}
%\end{table}
%\begin{table}[htp!]
%%\renewcommand{\arraystretch}{1.1}
%\centering
%\begin{tabular}{@{}p{2.5cm}p{1.2cm}p{1.2cm}p{1.2cm}p{1.2cm}p{1.2cm}@{}}\toprule
%Algorithm&\multicolumn{1}{l}{W\"ort.}&\multicolumn{1}{l}{Komm.}&\multicolumn{1}{l}{H\"oren}&\multicolumn{1}{l}{Lesen}&\multicolumn{1}{l}{Overall}\\ 
%\midrule
%TEM& .28 &	.82 &	.88 &	.79 &	.84 \\
%IBN& .71 &	.89 &	.83 &	.87 &	.90 \\
%TBN& .79 &	.91 &	.87 &	.89 &	.92 \\
%\bottomrule
%\end{tabular}
%\caption{Split-half reliability}
%\label{tab:halfsplit}
%\vspace{-0.8cm}
%\end{table}
%\begin{mdframed}[hidealllines=true,backgroundcolor=blue!20]
%In this view, we believe that our approach provides more reliable evaluations than the traditional one. Tab.~\ref{tab:halfsplit} shows the internal consistency of the three tests evaluated using the split-half methodology: the questions are randomly divided into two subset of equal size; the two half-tests thus obtained are used to generate two independent evaluations of the students; the correlation coefficient between the outputs of the two half-tests adjusted with the Spearman-Brown correction is retained as a measure of the test reliability. Results show that the reliability of both Bayesian approaches (IBN and TBN) is larger than that of the traditional method. TBN has the highest reliability, as it considers the correlation between different skills and thus all answers contribute to the determination of the level of a skill, even those referring to different skills. 
%Concerning the adaptive case, we have considered different thresholds for the entropy. The left plot in Fig.~\ref{fig:agreement} shows the relative agreement of the adaptive tests using IBN and TBN with the non-adaptive TBN, whereas the right plot shows the average number of questions asked to the students. Both models show a strong reduction in the number of questions as the threshold on the entropy increases. For instance, using the TBN model, the number of questions reduces of about 70\% passing from the non-adaptive test to an adaptive test with entropy threshold $\tilde{H}=0.5$. Yet, the test still provides the same evaluations as the non-adaptive one in more than 70\% of the cases. Such an accuracy loss might be problematic. However, even decreasing the entropy threshold to $\tilde{H}=0.19$, we can still save 20 question on average (to a maximum of 65 questions in the best case) at the price of a reduction of only 3\% in the accuracy. This shows that a significant number of questions provides very little information about the student level of knowledge. Our adaptive approach makes it possible to avoid wasting students' time asking uninformative questions. 
%\end{mdframed}
%\begin{figure}[htp!]
%\centering
%\begin{tikzpicture}
%\begin{axis}[title={Agreement},width=6cm, height=4cm,xlabel={Entropy Threshold $\tilde{H}$},xmin=0, xmax=0.5, ymin=0.7, ymax=1, xtick={0,.1,.2,.3,.4,.5}, ytick={.7,.8,.9,1}, legend pos=north east, ymajorgrids=true, xmajorgrids=true]
%\addplot[color=black,mark=square] coordinates { (0.0625,0.99833) (0.125,0.99217) (0.1875,0.9685) (0.25,0.89583) (0.375,0.75183) (0.5,0.74517)};    
%\addlegendentry{TBN}
%\addplot[color=black!60,mark=star] coordinates { (0.0625,0.94733)(0.125,0.946000)(0.1875,0.943333)(0.25,0.8965000)(0.375,0.744000)(0.5,0.724666)};
%\addlegendentry{IBN}
%\end{axis}
%\end{tikzpicture}
%\begin{tikzpicture}
%\begin{axis}[ title={Number of questions},width=6cm, height=4cm, xlabel={Entropy Threshold $\tilde{H}$}, xmin=0, xmax=0.5, ymin=0, ymax=100, xtick={0,.1,.2,.3,.4,.5}, ytick={0,20,40,60,80,100}, legend pos=north east, ymajorgrids=true, xmajorgrids=true]
%\addplot[color=black!60,mark=star] coordinates { (0.0625,89.741) (0.125,87.229) (0.1875,81.309) (0.25,66.745) (0.375,32.5) (0.5,29.001)};    
%\addlegendentry{IBN}
%\addplot[color=black,mark=square] coordinates { (0.0625,88.553) (0.125,83.476) (0.1875,75.071) (0.25,58.484) (0.375,30.722) (0.5,29)};
%\addlegendentry{TBN}
%\end{axis}
%\end{tikzpicture}
%\caption{Agreement between the adaptive methods and the non-adaptive TBN (left) and average number of questions asked by the adaptive methods (right)}\label{fig:agreement}
%\vspace{-.8cm}
%\end{figure}
%As providing precise values for these probabilities can be very hard even for expert teachers, the subjective opinion about the likelihood of an event was described by using the verbal-numerical probability scales in Table \ref{tb:labels}. 
%The question and skill CPTs are given in tables \ref{tb:questionCPTs} and \ref{tb:skillCPTs}.
%make the model specification complete, the CPTs of the skills, i.e., $\{ P(X_i|\mathrm{\Pi_{X_i}})\}_{i=1}^4$, and those of the questions, i.e., $\{ P(Y_j|B_{Y_j})\}_{j=1}^{95}$ should be quantified.
%In previous work \cite{mangili2016a}, these parameters have been learned from the available data. Results have shown that the resulting model can achieve larger internal consistency than the traditional evaluation method in assessing the student level. Given this result, a data-driven approach is used here to provide \emph{ground truth} evaluations for each student. This approach, hereafter referred to as EM-BN, uses expectation maximization (EM) to learn the CPTs of the BN model in  figure \ref{fig:fourskills} and infer the students level in the four skills. 
%The focus of the paper, however, is to provide an approach to adaptive testing entirely based on expert elicitation. Therefore, 
%As we are interested in adaptive testing entirely based on expert elicitation, teachers have been asked to elicit all CPTs in the model. 
%Following \cite{zaffalon2012evaluating}, the $u_{65}$ descriptors slightly increases the discounted accuracy contribution (i.e., $1/L$ is the set of labels includes the correct one and zero otherwise) with a convex xxx. Note that this descriptors is knows. 
%The above-described model has been used to evaluate the skills of the 451 students in the dataset. 
%As a ground truth we take the result based on the percentage of correct answers.
%Then, we have compared the effectiveness of the BN and CN-based adaptive approaches.
%?? traditional method, come nel paper ITS, EM?? \todo{scegliere ground truth}. 
% we adopt the following performance descriptors:
%\begin{itemize}
%\item discounted accuracy (disA), i.e., the average value of $\frac{I_{\mathcal{L}}(l)}{L}$, where $I$ is the indicator function and $l$ the true level. This can be regarded as the most unfavorable way (among the reasonable ones) to evaluate
%credal classifiers;
%\item set accuracy (setAcc), i.e., the average value of $I_{\mathcal{L}(l)}$. This can be regarded as the performance measure most favorable to a credal classifier;
%\item the utility-discounted accuracy U65 and U80 \cite{zaffalon2012evaluating}, i.e., the average value of a concave function $u\left(\frac{I_{\mathcal{L}}(l)}{L}\right)$ of the discounted utility. This measure assumes that being undecided between $k$ levels is better than choosing the correct level with probability $1/k$. This measure accounts for risk aversion, which increases with $u(0.5)$. We have $u(0.5)=0.65$ and $u(0.5)=0.8$ for, respectively the U65 and the U80 measures; \footnote{Notice that with $u(0.5)=0.5$ we obtain the discounted accuracy measure.}
%\item the average number of times the credal test is determinate, that is, it returns a single level (DET);
%\item the accuracy of the Bayesian test when the credal test is determinate (detA) and when it is indeterminate (indA). These indicators can show the capability of the credal test to identify the students which are more difficult to evaluate;
%\item the average cardinality of $\mathcal{L}$ (Csize).
%\end{itemize}

ISIPTA 2021 papers have to be prepared using \LaTeX.
To start writing your paper, copy \texttt{isipta2021-paper\_template.tex} and replace author and content with your own.

The ISIPTA 2021 paper style is based on the \textsf{jmlr} class with
the \texttt{pmlr,twocolumn} class options.\footnotemark
\footnotetext{%
  You can find the \textsf{jmlr} documentation at \url{https://ctan.org/pkg/jmlr}.
  The bundle you can download there also includes a sample file with more usage examples.
  Version 1.21 and before are known not to work correctly and produce wrong margins.
  Version 1.23 and after have been tested to work.
  The necessary files from version 1.27 are included in the author kit.
}
This class already loads some commonly used packages, so please have a look at its documentation in case you want to load extra packages.
Do not load packages that conflict with \textsf{jmlr}—you will get an error message—, as this would complicate creating the proceedings.

We have added a style file, \textsf{isipta2021} that sets up some more style details and loads some packages.
This includes font definitions, so do not load your own or change the font size or other style-related aspects.
Also do not change the line width, margins, line spacing and other vertical whitespace between layout elements (so do not use \verb|\vspace|).

You can load extra packages available in the big TeX distributions (Mik\TeX, \TeX{Live}, Mac\TeX) as long as your text compiles with PDF\LaTeX.
Please avoid using obsolete commands, such as \verb|\rm|, and obsolete packages, such as \textsf{epsfig}.\footnote{See \url{https://ctan.org/pkg/l2tabu}.}
\swap[ ]{in the header of your source file.}{But feel free to include your own macros}

\section{Practical Guidelines}
\subsection{Page Limits}
\emph{Full papers} have a length of four to nine pages.
This excludes the list of references and informational sections such as Acknowledgments.
There is the option to provide supplementary materials, such as videos, pictures, code, data-sets, etc.
However, these do not have to be considered by the reviewers, so do make sure the paper is self-contained without them.

\emph{Short papers} have a length of three pages.
Again, this excludes the list of references and informational sections such as Acknowledgments.
However, no supplementary material is accepted in this case.

Please be aware that these are the limits for the final versions of the papers. It is recommended that first submissions do not use all the allowed space, in order to leave some room to take into account reviewers suggestions.

\subsection{Submission \& Filenames}
Your ISIPTA 2021 paper must be submitted via the ISIPTA 2021 Easychair page: \url{https://easychair.org/conferences/?conf=isipta2021}

For the submission deadlines, please check the conference website: \url{http://www.sipta.org/isipta21}.

All the files you submit should follow the JMLR-required format: the surname of the first author, in lowercase, without spaces and accents, followed by \texttt{21} (for the publication year), and optional letter in case there are multiple submissions with the same first author surname, and then the extension.
So if this were a submission, the source file must be named \texttt{alandvonvereweg21.tex} and the compiled one \texttt{alandvonvereweg21.pdf} or, in case of multiple submissions, \texttt{alandvonvereweg21a.tex} and \texttt{alandvonvereweg21a.pdf}.
Supplementary materials gets a \texttt{-supp} affix to the name, so for this ‘submission’ a relevant dataset would, e.g., have to be named \texttt{alandvonvereweg21a-supp.csv}.

For the initial submission, you will have to submit the pdf of your paper and separately, if present, all the supplementary materials in one zip file.

If your paper has been accepted and you are preparing the version that will be published, download the copyright form from the conference web page.
Print, fill, and sign the form.
The name of the PDF scan should be the ID with the suffix \texttt{Permission} (for example, \texttt{alandvonvereweg21aPermission.pdf}).

For the final submission, you should update the paper pdf and, if supplementary materials changed, the supplementary materials zip file.
Additionally, you should create a zip file \texttt{alandvonvereweg21a.zip} with all the paper source files (\texttt{alandvonvereweg21a.tex}, \texttt{alandvonvereweg21a.bib}, etc.) and the permission pdf \texttt{alandvonvereweg21aPermission.pdf}.
This zip file should then be uploaded on Easychair as well.

\section{Sections}
Section titles must be in title case at all levels.

\subsection{Subsections}
\subsubsection{Subsubsections}
Avoid using subsubsections.
In case you need to structure your subsections more finely than in paragraphs, have a look at the unnumbered inline sectioning styles below.

\paragraph{‘Paragraphs’}
‘Paragraphs’ are produced using \verb|\paragraph|.
They can be used directly below sections or subsections.

You can have more than one paragraph in a \LaTeX\ ‘paragraph’.

\subparagraph{Subparagraphs}
Subparagraphs are produced using \verb|\subparagraph|.

\section{Cross-Referencing}
Always use \verb|\label| and \verb|\ref|—or one of the commands described below—when cross-referencing.
For example, the next section is Section~\ref{sec:math}.

The \textsf{jmlr} class provides some convenient cross-referencing commands: \verb|\sectionref|, \verb|\equationref|, \verb|\tableref|, \verb|\figureref|, \verb|\algorithmref|, \verb|\theoremref|, \verb|\lemmaref|, \verb|\remarkref|, \verb|\corollaryref|, \verb|\definitionref|, \verb|\conjectureref|, \verb|\axiomref|, \verb|\exampleref| and \verb|\appendixref|.
The argument of these commands may either be a single label or a comma-separated list of labels.
Examples: \sectionref{sec:math} or \sectionref{sec:intro,sec:math} or \sectionref{sec:intro,sec:math,sec:tables,sec:figures}.

\section{Math}\label{sec:math}
The \textsf{jmlr} class loads the \textsf{amsmath} package, so you can use any of the commands and environments defined there.\footnotemark
\footnotetext{%
  See the \textsf{amsmath} documentation at \url{https://ctan.org/pkg/amsmath} for further details.
}
Furthermore, the \textsf{isipta2021} style file loads the \textsf{mathtools} package, which extends \textsf{amsmath} with even more useful commands.\footnotetext{%
  See the \textsf{mathtools} documentation at \url{https://ctan.org/pkg/mathtools} for further details.
}

Use the \textsf{amsmath} environments for displayed equations.
So, specifically, use the \texttt{equation} environment instead of \verb|$$...$$| and the \texttt{align} environment instead of \texttt{eqnarray}.\footnote{For reasons why you shouldn't use the obsolete \texttt{eqnarray} environment, see Lars Madsen, \textit{Avoid eqnarray!} TUGboat 33(1):21--25, 2012.}
An \texttt{equation}:
\begin{equation}\label{eq:example}
  0 = 1 - 1
\end{equation}
Two \texttt{align}'ed equations:
\begin{align*} % no numbers with starred version
  1 + 2 &= 3,\\
  1 - 2 &= -1.
\end{align*}
Equations can also be put inline, of course.
For example, \equationref{eq:example}: \(0=1+1\). % $0=1+1$ also works

The \textsf{amsmath} and \textsf{mathtools} packages provide a lot of nice functionality, such as many common math operators, such as \(\sin\), \(\max\), etc. and also commands for defining new ones.

The following theorem-like environments are predefined by the \textsf{jmlr} class: \texttt{theorem}, \texttt{example}, \texttt{lemma}, \texttt{proposition}, \texttt{remark}, \texttt{corollary}, \texttt{definition}, \texttt{conjecture} and \texttt{axiom}.
You can use the \texttt{proof} environment to display the proof if need be, as in \theoremref{thm:example}.
\begin{theorem}[Example]\label{thm:example}
  All elements of the powerset of \(\varnothing\) are equal.
\end{theorem}
\begin{proof}\textbf{of \theoremref{thm:example}.} % You can leave off the reference in the \textbf if the proof comes right after that what is to be proven
  The powerset of \(\varnothing\) contains a unique element, \(\varnothing\).
\end{proof}

\section{Floats}\label{sec:floats}
Floats, such as figures, tables and algorithms, are moving objects and are supposed to float to the nearest convenient location.
Please do not force them to go in a particular place and do not
put the figure or table in the middle of a paragraph.
Floats are supposed to have a little extra space above and below them to make them stand out from the rest of the text.
This extra spacing is put in automatically and shouldn't need modifying.

\subsection{Figures}\label{sec:figures}
Figures should go in the \texttt{figure} environment.
Within this environment, use \verb|\floatconts| (defined by \textsf{jmlr}) to correctly position the caption and center the image.
To see how this works, look at the definition of \figureref{fig:example}.
Sub-figures can be created using \verb|\subfigure|, which is also defined by the \textsf{jmlr} class.
Use \verb|\includegraphics| for external graphics files but omit the file extension.
Supported formats are \textsf{pdf} (preferred for vector drawings and diagrams), \textsf{png} (preferred for screenshots), and \textsf{jpeg} (preferred for photographs).
Do not use \verb|\epsfig| or \verb|\psfig|.
If you want to scale the image, it's better to use a fraction of the line width rather than an explicit length.
For example, see \figureref{fig:page}.

\begin{figure}
 % The first argument is the label,
 % the caption goes in the second argument, and
 % the figure contents go in the third argument.
\floatconts
  {fig:example}
  {\caption{Example Image}}
  {
    \subfigure[A page]{%
      \label{fig:page}
      \includegraphics[width=0.5\linewidth,page=1]{isipta2021-poster-abstract_template}
    }
    \quad
    \subfigure[A coordinate system]{
      \label{fig:coords}%
      {
        \begin{tikzpicture}[xscale=1.5]
          \coordinate (origin);
          \draw[->] (origin) -- +(1cm,0) node[below] {$x$};
          \draw[->] (origin) -- +(0,1cm) node[left] {$y$};
          \fill[gray] (45:1cm) circle[radius=.2cm];
        \end{tikzpicture}
      }
    }
  }
\end{figure}

Don't use \verb|\graphicspath|.
If the images are contained in a subdirectory, specify this when you include the image, for example \verb|\includegraphics{figures/mypic}|.

\subsection{Tables}\label{sec:tables}
Tables should go in the \texttt{table} environment.
Within this environment use \verb|\floatconts| (defined by \textsf{jmlr})
to set the caption correctly and center the table contents.
To see how this works, look at the definition of \tableref{tab:example}.
Similarly to subfigures, subtables can be created using \verb|\subtable|, which is also defined by the \textsf{jmlr} class.

\begin{table}[htbp]
 % The first argument is the label,
 % the caption goes in the second argument, and
 % the table contents go in the third argument.
\floatconts{tab:example}%
  {\caption{An example table with nice horizontal lines and without unnecessary vertical lines}}%
  {
    \begin{tabular}{rl}
      \toprule % from booktabs package
      \bfseries Dataset & \bfseries Result\\
      \midrule % from booktabs package
      Data1 & 0.12345\\
      Data2 & 0.67890\\
      Data3 & 0.54321\\
      Data4 & 0.09876\\
      \bottomrule % from booktabs package
    \end{tabular}
  }
\end{table}

\subsection{Algorithms}\label{sec:algorithms}

Enumerated textual algorithms can be displayed using the \texttt{algorithm} environment; see the \textsf{jmlr} class documentation and sample file for a description.\footnotemark[1]
Code can be displayed using the automatically loaded \texttt{algorithm2e} environment defined by the \textsf{algorithm2e} package.\footnote{See the \textsf{algorithm2e} documentation at \url{https://ctan.org/pkg/algorithm2e}.}

\section{Citations and Bibliography}\label{sec:cite}
The references must be managed using Bib\TeX.
The \textsf{jmlr} class automatically loads \textsf{natbib} and the appropriate style file.
So you have to use the commands provided by this package.\footnote{You can find the \textsf{natbib} documentation at \url{https://ctan.org/pkg/natbib}.}
This template file has the citations defined in the accompanying BibTeX file \texttt{isipta2021-template.bib}.
For a parenthetical citation use \verb|\citep|.
For example: “Good introductions to the topic are available \citep{latexcompanion}.”
For a textual citation use \verb|\citet|.
For example: “\citet{einstein} discusses electrodynamics of moving bodies.”
The citation style can be chosen using the \verb|\citestyle| command in the preamble.

The bibliography is displayed using \verb|\bibliography|.

\appendix

\section{Boring Stuff}\label{apd:first}
In case you wish to structure your paper using appendices, you can do so.
However, they also count towards the page limit.

\textbf{But, everything below this line does not count towards the page limit.}
%% For papers, sections below this line do not count towards the page limit

\acks{%
Acknowledgments can go here.
This includes acknowledgments for financial support as well as for support from other people.
}

\section*{Author Contributions}
If you wish, you can mention which of the authors did what.
Otherwise this section can be omitted.

\section*{Competing Interests}
In case any author has some competing interest it has to be mentioned here.
Otherwise this section can be omitted.

\bibliography{isipta2021-template}
% should be alandvonvereweg21.bib if this were a submission

\end{document}